Mixed effect tweedie model in TMB

### 2019-04-08 analysis/tmb-sensor-explore/2-fit-temperature.Rmd
prelim comparison withheld data with model estimates... found error in how predict function deals with newdata

```{r}
# FIXME: Error in `$<-.data.frame`(`*tmp*`, "est", value = c(7.12313541783276, :replacement has 4716 rows, data has 4717
predictions <- predict(m_temp1)

attributes(m_temp1)

# find source of error in predict function
object <- m_temp1

 nd <- object$data # this has 3,774 rows of data
 lp <- object$tmb_obj$env$last.par # this has 6583 length(lp)
  r <- object$tmb_obj$report(lp)
    length(r$omega_s) # 386
    length(r$epsilon_st) # this is 6176
    length(r$b_rw_t) # this has 16 time steps
    length(r$eta_i) # this is a vector of 3773

    nd$est <- r$eta_i # ERROR OCCURS HERE
    nd$est_fe <- r$eta_fixed_i # ERROR OCCURS HERE

attributes(r)

@@

```{r}
plot_map(predictions$data, "est_fe") +
  ggtitle("Prediction (fixed effects only)") +
  scale_colour_viridis_c()

plot_map(predictions$data, "est_re_s") +
  ggtitle("Spatial random effects only") +
  scale_colour_gradient2()

plot_map(predictions$data, "est_re_st") +
  ggtitle("Spatiotemporal random effects only") +
  scale_colour_gradient2()

ggplot(predictions$data, aes(est, temperature_c)) +
  geom_point() +
  facet_wrap(~year) +
  coord_fixed() +
  geom_abline()

ggplot(predictions$data, aes(est, residuals)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~year)

ggplot(predictions$data, aes(depth_scaled, residuals)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~year)

ggplot(predictions$data, aes(depth_scaled, residuals)) +
  geom_point() +
  geom_smooth()

ggplot(predictions$data, aes(X, Y, colour = residuals)) +
  geom_point() +
  facet_wrap(~year) +
  coord_fixed() +
  scale_color_gradient2()

# Looks a bit curved, let's try with a quadratic

m_temp2 <- sdmTMB(d_fit, temperature_c ~ 0 + as.factor(year) + poly(depth_scaled, 2),
  time = "year", spde = spde, family = gaussian(link = "identity"),
  silent = TRUE)

predictions2 <- predict(m_temp2)
predictions2$data$residuals <- residuals(m_temp2)

ggplot(predictions2$data, aes(depth_scaled, residuals)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~year)

ggplot(predictions2$data, aes(depth_scaled, residuals)) +
  geom_point() +
  geom_smooth()

ggplot(predictions2$data, aes(X, Y, colour = residuals)) +
  geom_point() +
  facet_wrap(~year) +
  coord_fixed() +
  scale_color_gradient2()

# Try with a third degree polynomial:
m_temp3 <- sdmTMB(d_fit, 
  temperature_c ~ 0 + as.factor(year) + poly(depth_scaled, 3),
  time = "year", spde = spde, family = gaussian(link = "identity"),
  silent = FALSE)

m_temp3_rw <- sdmTMB(d_fit,
  temperature_c ~ 0 + as.factor(year),
  time_varying =  ~ 0 + poly(depth_scaled, 3),
  include_spatial = FALSE, ar1_fields = TRUE,
  time = "year", spde = spde, family = gaussian(link = "identity"),
  silent = FALSE)

predictions3 <- predict(m_temp3)
predictions3$data$residuals <- residuals(m_temp3)

predictions3_rw <- predict(m_temp3_rw)
predictions3_rw$data$residuals <- residuals(m_temp3_rw)

ggplot(predictions3_rw$data, aes(depth_scaled, residuals)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~year)

ggplot(predictions3$data, aes(depth_scaled, residuals)) +
  geom_point() +
  geom_smooth()

ggplot(predictions3$data, aes(X, Y, colour = residuals)) +
  geom_point() +
  facet_wrap(~year) +
  coord_fixed() +
  scale_color_gradient2()

@@

d_fit$log_temperature_c <- log(d_fit$temperature_c)
m_temp4 <- sdmTMB(d_fit, log_temperature_c ~ 0 + as.factor(year) + poly(depth_scaled, 3),
  time = "year", spde = spde, family = gaussian(link = "identity"),
  silent = TRUE)

predictions4 <- predict(m_temp4)
predictions4$data$residuals <- residuals(m_temp4)

ggplot(predictions4$data, aes(depth_scaled, residuals)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~year)

ggplot(predictions4$data, aes(depth_scaled, residuals)) +
  geom_point() +
  geom_smooth()

ggplot(predictions4$data, aes(X, Y, colour = residuals)) +
  geom_point() +
  facet_wrap(~year) +
  coord_fixed() +
  scale_color_gradient2()

@@

m_temp5 <- sdmTMB(d_fit, temperature_c ~ 0 + as.factor(year) + poly(depth_m, 3),
  time = "year", spde = spde, family = gaussian(link = "identity"),
  silent = TRUE)

predictions5 <- predict(m_temp5)
predictions5$data$residuals <- residuals(m_temp5)

ggplot(predictions5$data, aes(depth_scaled, residuals)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~year)

ggplot(predictions5$data, aes(depth_scaled, residuals)) +
  geom_point() +
  geom_smooth()

@@

ggplot(predictions5$data, aes(est, residuals)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~year)

ggplot(predictions5$data, aes(est, residuals)) +
  geom_point() +
  geom_smooth()

aic <- function(m) {
  k <- length(m$model$par)
  nll <- m$model$objective
  2 * k - 2 * (-nll)
}
aic(m_temp1)
aic(m_temp2)
aic(m_temp3)
aic(m_temp4)
aic(m_temp5)

@@

ggplot(predictions3$data, aes(est, temperature_c)) +
  geom_point() +
  facet_wrap(~year) +
  coord_fixed() +
  geom_abline()

# R2
cor(predictions3$data$est, predictions3$data$temperature_c)^2

@@

predictions3_nd <- predict(m_temp3, newdata = sdmTMB::qcs_grid)
saveRDS(predictions3_nd$data, file = "analysis/tmb-sensor-explore/qcs-temp.rds")

ggplot(predictions3$data, aes(X, Y, colour = residuals)) +
  geom_point() +
  facet_wrap(~year) +
  coord_fixed() +
  scale_color_gradient2()

# A short function for plotting our predictions:
plot_map <- function(dat, column = "est") {
  ggplot(dat, aes_string("X", "Y", fill = column)) +
    geom_raster() +
    facet_wrap(~year) +
    coord_fixed()
}

plot_map(predictions3_nd$data, "est") +
  scale_fill_viridis_c(trans = "sqrt", option = "C") +
  ggtitle("Prediction (fixed effects + all random effects)")

plot_map(predictions3_nd$data, "est_fe") +
  ggtitle("Prediction (fixed effects only)") +
  scale_fill_viridis_c(trans = "sqrt")

plot_map(predictions3_nd$data, "est_re_s") +
  ggtitle("Spatial random effects only") +
  scale_fill_gradient2()

plot_map(predictions3_nd$data, "est_re_st") +
  ggtitle("Spatiotemporal random effects only") +
  scale_fill_gradient2()

temp_cog <- get_cog(predictions3_nd)

temp_cog %>% reshape2::dcast(year ~ coord, value.var = "est") %>%
  ggplot(aes(X, Y, colour = year)) + geom_path(arrow = arrow()) +
  scale_color_viridis_c()

temp_cog %>%
  ggplot(aes(year, est, ymin = lwr, ymax = upr)) +
  geom_line() +
  geom_ribbon(alpha = 0.2) +
  facet_wrap(~coord, scales = "free_y")
```

>>>>

```{r}
predictions <- predict(m_temp1)
```

@@

```{r}
plot_map(predictions$data, "est_fe") +
  ggtitle("Prediction (fixed effects only)") +
  scale_colour_viridis_c()
```

```{r}
plot_map(predictions$data, "est_re_s") +
  ggtitle("Spatial random effects only") +
  scale_colour_gradient2()
```

```{r}
plot_map(predictions$data, "est_re_st") +
  ggtitle("Spatiotemporal random effects only") +
  scale_colour_gradient2()
```

```{r}
ggplot(predictions$data, aes(est, temperature_c)) +
  geom_point() +
  facet_wrap(~year) +
  coord_fixed() +
  geom_abline()
```

```{r}
ggplot(predictions$data, aes(est, residuals)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~year)
```

```{r}
ggplot(predictions$data, aes(depth_scaled, residuals)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~year)
```

```{r}
ggplot(predictions$data, aes(depth_scaled, residuals)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~year)
```

```{r}
ggplot(predictions$data, aes(depth_scaled, residuals)) +
  geom_point() +
  geom_smooth()
```

```{r}
ggplot(predictions$data, aes(X, Y, colour = residuals)) +
  geom_point() +
  facet_wrap(~year) +
  coord_fixed() +
  scale_color_gradient2()
```

 Looks a bit curved, let's try with a quadratic

```{r}
m_temp2 <- sdmTMB(d_fit, temperature_c ~ 0 + as.factor(year) + poly(depth_scaled, 2),
  time = "year", spde = spde, family = gaussian(link = "identity"),
  silent = TRUE)

predictions2 <- predict(m_temp2)
predictions2$data$residuals <- residuals(m_temp2)
```

```{r}
ggplot(predictions2$data, aes(depth_scaled, residuals)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~year)
```

```{r}
ggplot(predictions2$data, aes(depth_scaled, residuals)) +
  geom_point() +
  geom_smooth()
```

```{r}
ggplot(predictions2$data, aes(X, Y, colour = residuals)) +
  geom_point() +
  facet_wrap(~year) +
  coord_fixed() +
  scale_color_gradient2()
```

Try with a third degree polynomial:

```{r}
m_temp3 <- sdmTMB(d_fit, 
  temperature_c ~ 0 + as.factor(year) + poly(depth_scaled, 3),
  time = "year", spde = spde, family = gaussian(link = "identity"),
  silent = FALSE)

predictions3 <- predict(m_temp3)
predictions3$data$residuals <- residuals(m_temp3)
```

```{r}
m_temp3_rw <- sdmTMB(d_fit,
  temperature_c ~ 0 + as.factor(year),
  time_varying =  ~ 0 + poly(depth_scaled, 3),
  include_spatial = FALSE, ar1_fields = TRUE,
  time = "year", spde = spde, family = gaussian(link = "identity"),
  silent = FALSE)

predictions3_rw <- predict(m_temp3_rw)
predictions3_rw$data$residuals <- residuals(m_temp3_rw)
```

```{r}
ggplot(predictions3_rw$data, aes(depth_scaled, residuals)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~year)
```

```{r}
ggplot(predictions3$data, aes(depth_scaled, residuals)) +
  geom_point() +
  geom_smooth()
```

```{r}
ggplot(predictions3$data, aes(X, Y, colour = residuals)) +
  geom_point() +
  facet_wrap(~year) +
  coord_fixed() +
  scale_color_gradient2()
```

@@

```{r}
d_fit$log_temperature_c <- log(d_fit$temperature_c)
m_temp4 <- sdmTMB(d_fit, log_temperature_c ~ 0 + as.factor(year) + poly(depth_scaled, 3),
  time = "year", spde = spde, family = gaussian(link = "identity"),
  silent = TRUE)

predictions4 <- predict(m_temp4)
predictions4$data$residuals <- residuals(m_temp4)
```

```{r}
ggplot(predictions4$data, aes(depth_scaled, residuals)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~year)
```

```{r}
ggplot(predictions4$data, aes(depth_scaled, residuals)) +
  geom_point() +
  geom_smooth()
```

```{r}
ggplot(predictions4$data, aes(X, Y, colour = residuals)) +
  geom_point() +
  facet_wrap(~year) +
  coord_fixed() +
  scale_color_gradient2()
```

@@

```{r}
m_temp5 <- sdmTMB(d_fit, temperature_c ~ 0 + as.factor(year) + poly(depth_m, 3),
  time = "year", spde = spde, family = gaussian(link = "identity"),
  silent = TRUE)

predictions5 <- predict(m_temp5)
predictions5$data$residuals <- residuals(m_temp5)
```

```{r}
ggplot(predictions5$data, aes(depth_scaled, residuals)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~year)
```

```{r}
ggplot(predictions5$data, aes(depth_scaled, residuals)) +
  geom_point() +
  geom_smooth()
```

@@

```{r}
ggplot(predictions5$data, aes(est, residuals)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~year)
```

```{r}
ggplot(predictions5$data, aes(est, residuals)) +
  geom_point() +
  geom_smooth()
```

```{r}
aic <- function(m) {
  k <- length(m$model$par)
  nll <- m$model$objective
  2 * k - 2 * (-nll)
}
aic(m_temp1)
aic(m_temp2)
aic(m_temp3)
aic(m_temp4)
aic(m_temp5)
```

@@

```{r}
ggplot(predictions3$data, aes(est, temperature_c)) +
  geom_point() +
  facet_wrap(~year) +
  coord_fixed() +
  geom_abline()
```

# R2

```{r}
cor(predictions3$data$est, predictions3$data$temperature_c)^2
```

@@

```{r}
predictions3_nd <- predict(m_temp3, newdata = sdmTMB::qcs_grid)
saveRDS(predictions3_nd$data, file = "qcs-temp.rds")
```

```{r}
ggplot(predictions3$data, aes(X, Y, colour = residuals)) +
  geom_point() +
  facet_wrap(~year) +
  coord_fixed() +
  scale_color_gradient2()
```

# A short function for plotting our predictions:

```{r}
plot_map <- function(dat, column = "est") {
  ggplot(dat, aes_string("X", "Y", fill = column)) +
    geom_raster() +
    facet_wrap(~year) +
    coord_fixed()
}
```

```{r}
plot_map(predictions3_nd$data, "est") +
  scale_fill_viridis_c(trans = "sqrt", option = "C") +
  ggtitle("Prediction (fixed effects + all random effects)")
```

```{r}
plot_map(predictions3_nd$data, "est_fe") +
  ggtitle("Prediction (fixed effects only)") +
  scale_fill_viridis_c(trans = "sqrt")
```

```{r}
plot_map(predictions3_nd$data, "est_re_s") +
  ggtitle("Spatial random effects only") +
  scale_fill_gradient2()
```

```{r}
plot_map(predictions3_nd$data, "est_re_st") +
  ggtitle("Spatiotemporal random effects only") +
  scale_fill_gradient2()
```

```{r}
m_temp3$model$par
# 2 * plogis(m_temp1$model$par[["ar1_phi"]]) - 1
exp(m_temp3$model$par[["ln_phi"]])

p_withheld <- predict(m_temp3, newdata = d_withheld)

#FIXME: p_withheld$data$temperature_c is all zeros?!?!?! 
# temp work around (first make sure both df in same order)
arrange(p_withheld$data, fishing_event_id) 
arrange(d_withheld, fishing_event_id) 
p_withheld$data$temperature_c <- d_withheld$temperature_c
```

```{r}
ggplot(p_withheld$data, aes(est, temperature_c)) +
  geom_point() +
  facet_wrap(~year) +
  coord_fixed() +
  xlim(3,12) +
  ylim(3,12) +
  geom_abline()
```

```{r}
temp_cog <- get_cog(predictions3_nd)

temp_cog %>% reshape2::dcast(year ~ coord, value.var = "est") %>%
  ggplot(aes(X, Y, colour = year)) + geom_path(arrow = arrow()) +
  scale_color_viridis_c()
```

```{r}
temp_cog %>%
  ggplot(aes(year, est, ymin = lwr, ymax = upr)) +
  geom_line() +
  geom_ribbon(alpha = 0.2) +
  facet_wrap(~coord, scales = "free_y")
```
